---
title: "Arcara 2024 - Regression Method<br>for normative data - Tutorial"
output:
  html_document: default
  word_document: default
date: "2024-03-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Introduction

This tutorials explain how to use the code for applying the method in *Arcara G.(2024) Improving Equivalent scores: a new method for regression model selection.* 

## Set up R
To work with the method you need to load some functions that will be used. Here I included all the required functions in a folder called `R functions` (these will be wrapped in an R package in the future).

```{r wd, echo=FALSE}
setwd( "/Users/giorgioarcara/Documents/Lavori Personali/Improving ES/Reg-Method-Norm-Data-CodeApp/")
```

```{r functions}
source("R_functions/adjscores_A2024_v3_mod.R")
source("R_functions/formula_transf_text.R")
source("R_functions/model_transf_text.R")
source("R_functions/transf_functions.R")
source("R_functions/tolLimits.adjscores.R")
source("R_functions/tolLimits.obs.R")
source("R_functions/ES.R")

```

I then load some required packages. 

```{r packages, results="markup"}
require(effects)
require(car)
require(performance)
```

## Import data

In the following lines I import the data and make some check everything is ok

```{r pressure, echo=TRUE}
Test.dat = read.csv("Original_Data/MOCA_Dataset.csv", sep=",", dec=".")

dim(Test.dat)

head(Test.dat)

str(Test.dat)

# fix values for participants with zero Education otherwise some transformations (e.g. 1/x, log) could give issues.
Test.dat[Test.dat$Education==0, "Education"] = 1

Test.dat = na.omit(Test.dat)

```

## Use the method to select regression model

```{r method}
Test.ARC.res = adjscores_A2024(df = Test.dat, dep="Score", 
                             age="Age", edu="Education", sex="Sex",
                             dep.range = c(0, 100))
```

Inspect model results. The results is a list of objects (see `adjscores_A2024.R` file for details):

* the first is a dataset in which a new column with adjusted scores is added

```{r dataset}
head(Test.ARC.res$new.df)
```

* the second is the linear model estimated (here I use `summary` to better inspect the model) and I plot the partial effects

```{r lm}
summary(Test.ARC.res$lm.model)
plot(allEffects(Test.ARC.res[[2]], partial.residuals=TRUE))
``` 


* I can also print the final formula
```{r formula}
print(Test.ARC.res$model_text)
```

# Model diagnositc
An appropriate model selection should also include some diagnostics
```{r diagnostic}
plot(check_normality(Test.ARC.res$lm.model))
plot(check_heteroscedasticity(Test.ARC.res$lm.model))
# some disomogeneity in variance, but overall a good fit
```


# Calculate Equivalent Scores

To calculate thresholds for Equivalent Scores I can use some some few additional code.

```{r ES}
Test.ES = ES(adjscores=Test.ARC.res$new.df$ADJ_SCORES)

print(Test.ES)
```

# Citing the Method
If you use the regression method script please cite: 
*Arcara G. (2024) Improving Equivalent Scores: A new method for regression model selection*

If you use the ES, please cite: 
*Aiello, E. N., & Depaoli, E. G. (2022). Norms and standardizations in neuropsychology via equivalent scores: software solutions and practical guides. Neurological Sciences, 43(2), 961-966.*

